---
title: "Introduction with R"
author: "Jobb Rodriguez"
output: html_document
---
# Intro to basics

## How it works
R uses the \# sign to add comments.

**Exercise**
``` {r}
# Calculate 3 + 4
3 + 4
# Calculate 6 + 12
6 + 12
```

## Arithemetic with R
In its most basic form, R can be used as a simple calculator. Consider the following arithmetic operators:

* Addition: +
* Subtraction: -
* Multiplication: *
* Division: /
* Exponentiation: ^
* Modulo: %%

The last two might need some explaining:

* The ^ operator raises the number to its left to the power of the number to its right: for example 3^2 is 9.
* The modulo returns the remainder of the division of the number to the left by the number on its right, for example 5 modulo 3 or 5 %% 3 is 2.

**Exercise**
```{r}
# An addition
5 + 5 

# A subtraction
5 - 5 

# A multiplication
3 * 5

 # A division
(5 + 5) / 2 

# Exponentiation
2 ^ 5

# Modulo
28 %% 6
```

## Variable assignment
A basic concept in (statistical) programming is called a **variable**.

A variable allows you to store a value (e.g. 4) or an object (e.g. a function description) in R. You can then later use this variable's name to easily access the value or the object that is stored within this variable.

You can assign a value 4 to a variable my_var with the command
```{r}
my_var <- 4
```

**Exercise**
```{r}
# Assign the value 42 to x
x <- 42

# Print out the value of the variable x
x

# Assign the value 5 to the variable my_apples
my_apples <- 5

# Print out the value of the variable my_apples
my_apples

# Assign a value to the variables my_apples and my_oranges
my_apples <- 5
my_oranges <- 6

# Add these two variables together
my_apples + my_oranges

# Create the variable my_fruit
my_fruit <- my_apples + my_oranges

# Print the value of the variable my_fruit
my_fruit
```

## Basic data types in R
R works with numerous data types. Some of the most basic types to get started are:

* Decimal values like ```4.5``` are called **numerics**.
* Whole numbers like ```4``` are called **integers**. Integers are also numerics.
* Boolean values (```TRUE``` or ```FALSE```) are called **logical**.
* Text (or string) values are called **characters**.

Note how the quotation marks in the editor indicate that ```"some text"``` is a string.

**Exercise**
```{r}
# Change my_numeric to be 42
my_numeric <- 42

# Change my_character to be "universe"
my_character <- "universe"

# Change my_logical to be FALSE
my_logical <- FALSE
```

## What's that data type?
You can use the ```class()``` function to check the data type of a variable.

**Exercise**
```{r}
# Declare variables of different types
my_numeric <- 42
my_character <- "universe"
my_logical <- FALSE 

# Check class of my_numeric
class(my_numeric)

# Check class of my_character
class(my_character)

# Check class of my_logical
class(my_logical)
```