---
title: "Introduction with R"
author: "Jobb Rodriguez"
output: html_document
---
# Matrices

## What's a matrix?
In R, a matrix is a collection of elements of the same data type (numeric, character, or logical) arranged into a fixed number of rows and columns. Since you are only working with rows and columns, a matrix is called two-dimensional.

You can construct a matrix in R with the ```matrix()``` function. Consider the following example:

```matrix(1:9, byrow = TRUE, nrow = 3)```

In the ```matrix()``` function:

* The first argument is the collection of elements that R will arrange into the rows and columns of the matrix. Here, we use ```1:9``` which is a shortcut for ```c(1, 2, 3, 4, 5, 6, 7, 8, 9)```.
* The argument byrow indicates that the matrix is filled by the rows. If we want the matrix to be filled by the columns, we just place ```byrow = FALSE```.
* The third argument ```nrow``` indicates that the matrix should have three rows.

**Exercise**

Construct a matrix with 3 rows that contain the numbers 1 up to 9
```{r echo = FALSE}
matrix(1:9, byrow = TRUE, nrow = 3)
```

## Analyze matrices, you shall
```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Create box_office
box_office <- c(new_hope, empire_strikes, return_jedi)

# Construct star_wars_matrix
star_wars_matrix <- matrix(box_office, byrow = TRUE, nrow = 3)
```

## Naming a matrix
Similar to vectors, you can add names for the rows and the columns of a matrix

```
rownames(my_matrix) <- row_names_vector
colnames(my_matrix) <- col_names_vector
```

We went ahead and prepared two vectors for you: ```region```, and ```titles```. You will need these vectors to name the columns and rows of ```star_wars_matrix```, respectively.

**Exericse**
```{r echo = FALSE}
# Vectors region and titles, used for naming
region <- c("US", "non-US")
titles <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")

# Name the columns with region
colnames(star_wars_matrix) <- region

# Name the rows with titles
rownames(star_wars_matrix) <- titles

# Print out star_wars_matrix
star_wars_matrix
```

## Calculating the worldwide box office
To calculate the total box office revenue for the three Star Wars movies, you have to take the sum of the US revenue column and the non-US revenue column.

In R, the function ```rowSums()``` conveniently calculates the totals for each row of a matrix. This function creates a new vector:

```rowSums(my_matrix)```

**Exercise**
```{r}
# Construct star_wars_matrix
star_wars_matrix <- matrix(box_office, 
                    nrow = 3, 
                    byrow = TRUE, 
                    dimnames = list(titles, region))

worldwide_vector <- rowSums(star_wars_matrix)
worldwide_vector
```

## Adding a column for the Worldwide box office


## Adding a row


## The total box office revenue for the entire sage


## Selection of matrix elements


## A little arithmetic with matrices