---
title: "Introduction with R"
author: "Jobb Rodriguez"
output: html_document
---
# Matrices

## What's a matrix?
In R, a matrix is a collection of elements of the same data type (numeric, character, or logical) arranged into a fixed number of rows and columns. Since you are only working with rows and columns, a matrix is called two-dimensional.

You can construct a matrix in R with the ```matrix()``` function. Consider the following example:

```matrix(1:9, byrow = TRUE, nrow = 3)```

In the ```matrix()``` function:

* The first argument is the collection of elements that R will arrange into the rows and columns of the matrix. Here, we use ```1:9``` which is a shortcut for ```c(1, 2, 3, 4, 5, 6, 7, 8, 9)```.
* The argument byrow indicates that the matrix is filled by the rows. If we want the matrix to be filled by the columns, we just place ```byrow = FALSE```.
* The third argument ```nrow``` indicates that the matrix should have three rows.

**Exercise**

Construct a matrix with 3 rows that contain the numbers 1 up to 9
```{r echo = FALSE}
matrix(1:9, byrow = TRUE, nrow = 3)
```

## Analyze matrices, you shall
```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Create box_office
box_office <- c(new_hope, empire_strikes, return_jedi)

# Construct star_wars_matrix
star_wars_matrix <- matrix(box_office, byrow = TRUE, nrow = 3)
```

## Naming a matrix
Similar to vectors, you can add names for the rows and the columns of a matrix

```
rownames(my_matrix) <- row_names_vector
colnames(my_matrix) <- col_names_vector
```

We went ahead and prepared two vectors for you: ```region```, and ```titles```. You will need these vectors to name the columns and rows of ```star_wars_matrix```, respectively.

**Exericse**
```{r echo = FALSE}
# Vectors region and titles, used for naming
region <- c("US", "non-US")
titles <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")

# Name the columns with region
colnames(star_wars_matrix) <- region

# Name the rows with titles
rownames(star_wars_matrix) <- titles

# Print out star_wars_matrix
star_wars_matrix
```

## Calculating the worldwide box office
To calculate the total box office revenue for the three Star Wars movies, you have to take the sum of the US revenue column and the non-US revenue column.

In R, the function ```rowSums()``` conveniently calculates the totals for each row of a matrix. This function creates a new vector:

```rowSums(my_matrix)```

**Exercise**
```{r}
# Construct star_wars_matrix
star_wars_matrix <- matrix(box_office, 
                    nrow = 3, 
                    byrow = TRUE, 
                    dimnames = list(titles, region))

worldwide_vector <- rowSums(star_wars_matrix)
worldwide_vector
```

## Adding a column for the Worldwide box office
You can add a column or multiple columns to a matrix with the ```cbind()``` function, which merges matrices and/or vectors together by column. For example:

```big_matrix <- cbind(matrix1, matrix2, vector1 ...)```


```{r}
# The worldwide box office figures
worldwide_vector <- rowSums(star_wars_matrix)

# Bind the new variable worldwide_vector as a column to star_wars_matrix
all_wars_matrix <- cbind(star_wars_matrix, worldwide_vector)
```

## Adding a row
```
# star_wars_matrix and star_wars_matrix2 are available in your workspace
star_wars_matrix  
star_wars_matrix2 

# Combine both Star Wars trilogies in one matrix
all_wars_matrix <- rbind(star_wars_matrix, star_wars_matrix2)
```

## The total box office revenue for the entire sage
```
# all_wars_matrix is available in your workspace
all_wars_matrix

# Total revenue for US and non-US
total_revenue_vector <- colSums(all_wars_matrix) 
  
# Print out total_revenue_vector
total_revenue_vector
```

## Selection of matrix elements
Similar to vectors, you can use the square brackets ```[ ]``` to select one or multiple elements from a matrix. Whereas vectors have one dimension, matrices have two dimensions. You should therefore use a comma to separate the rows you want to select from the columns. For example:

```my_matrix[1,2]``` selects the element at the first row and second column.
```my_matrix[1:3,2:4]``` results in a matrix with the data on the rows 1, 2, 3 and columns 2, 3, 4.
If you want to select all elements of a row or a column, no number is needed before or after the comma, respectively:

```my_matrix[,1]``` selects all elements of the first column.
```my_matrix[1,]``` selects all elements of the first row.

**Exercise**
```
# all_wars_matrix is available in your workspace
all_wars_matrix

# Select the non-US revenue for all movies
non_us_all <- all_wars_matrix[, 2]
  
# Average non-US revenue
mean(non_us_all)
  
# Select the non-US revenue for first two movies
non_us_some <- non_us_all[1:2]
  
# Average non-US revenue for first two movies
mean(non_us_some)
```

## A little arithmetic with matrices
Similar to what you have learned with vectors, the standard operators like ```+```, ```-```, ```/```, ```*```, etc. work in an element-wise way on matrices in R.

For example, ```2 * my_matrix``` multiplies each element of ```my_matrix``` by two.

As a newly-hired data analyst for Lucasfilm, it is your job to find out how many visitors went to each movie for each geographical area. You already have the total revenue figures in ```all_wars_matrix```. Assume that the price of a ticket was 5 dollars. Simply dividing the box office numbers by this ticket price gives you the number of visitors.

**Exercise**
```
# all_wars_matrix is available in your workspace
all_wars_matrix

# Estimate the visitors
visitors <- all_wars_matrix / 5
  
# Print the estimate to the console
visitors
```

Just like ```2 * my_matrix``` multiplied every element of my_matrix by two, ```my_matrix1 * my_matrix2``` creates a matrix where each element is the product of the corresponding elements in my_matrix1 and my_matrix2.

After looking at the result of the previous exercise, big boss Lucas points out that the ticket prices went up over time. He asks to redo the analysis based on the prices you can find in ```ticket_prices_matrix``` (source: imagination).

Those who are familiar with matrices should note that this is not the standard matrix multiplication for which you should use ```%*%``` in R.

**Exercise**
```
# all_wars_matrix and ticket_prices_matrix are available in your workspace
all_wars_matrix
ticket_prices_matrix

# Estimated number of visitors
visitors <- all_wars_matrix / ticket_prices_matrix

# US visitors
us_visitors <- visitors[, 1]

# Average number of US visitors
mean(us_visitors)
```